---
title: "EDAfile by Pranav Nair in ST558"
format: html
editor: visual
---

### Introduction

In this document, we will do exploratory data analysis of our data set. This data set shows health indicators that could show diabetes.  

The first step is to run the libraries necessary.

```{r}
library(readr)
library(tidyverse)
```


```{r}
diabetes_data <- read_csv("diabetes_012_health_indicators_BRFSS2015.csv", col_names = TRUE)
str(diabetes_data)
print(diabetes_data)


```

We see all these variables as numeric.  However, many of them should actually be factor variables with meaningful level names.

```{r}
diabetes_data_fixed <- diabetes_data |>
  mutate(Diabetes_012 = factor(Diabetes_012, levels = c(0,1,2), labels = c("No Diabetes", "Prediabetes", "Diabetes")),
         HighBP = factor(HighBP, levels = c(0,1), labels = c("No", "Yes")),
         HighChol = factor(HighChol, levels = c(0,1), labels = c("No", "Yes")),
         CholCheck = factor(CholCheck, levels = c(0,1), labels = c("No", "Yes")),
         Smoker = factor(Smoker, levels = c(0,1), labels = c("No", "Yes")),
         Stroke = factor(Stroke, levels = c(0,1), labels = c("No", "Yes")),
         HeartDiseaseorAttack = factor(HeartDiseaseorAttack, levels = c(0,1), labels = c("No", "Yes")),
         PhysActivity = factor(PhysActivity, levels = c(0,1), labels = c("No", "Yes")),
         Fruits = factor(Fruits, levels = c(0,1), labels = c("No", "Yes")),
         Veggies = factor(Veggies, levels = c(0,1), labels = c("No", "Yes")),
         HvyAlcoholConsump = factor(HvyAlcoholConsump, levels = c(0,1), labels = c("No", "Yes")),
         AnyHealthcare = factor(AnyHealthcare, levels = c(0,1), labels = c("No", "Yes")),
         NoDocbcCost = factor(NoDocbcCost, levels = c(0,1), labels = c("No", "Yes")),
         DiffWalk = factor(DiffWalk, levels = c(0,1), labels = c("No", "Yes")),
         Sex = factor(Sex, levels = c(0,1), labels = c("Female", "Male")),
         Age = factor(Age, levels = 1:13, labels = c("18-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+")),
         Education = factor(Education, levels = 1:6, labels = c("No Schooling", "Elementary", "Some High School", "High School Grad", "Some College", "College Grad")),
         Income = factor(Income, levels = 1:8, labels = c("<$10,000", "$10,000-$15,000", "$15,000-$20,000", "$20,000-$25,000", "$25,000-$35,000", "$35,000-$50,000", "$50,000-$75,000", ">$75,000"))
  )

print(diabetes_data_fixed)
```

Now let's check the data for missingness.
```{r}
sum(is.na(diabetes_data_fixed))
```

No missing data!  Now we can move on to the summarizations.

